The Creation of a Programming Language

### 2022-10-19 paper.tex
continue paper

\documentclass[12pt,a4paper,man]{apa7}
\usepackage{listings}
\usepackage{setspace}
\usepackage{hyperref}
\usepackage{dirtree}
\usepackage{catchfile}
\usepackage{times} % Use Times New Roman font
\usepackage[style=apa]{biblatex}

@@

\author{Florian Donnelly}
\title{The Creation of a Programming Language}
\authorsaffiliations{Gymnasium Burgdorf}
\shorttitle{Creation of a Programming Language}
\abstract{
    This is an abstract...
}

@@

@@

@@

\onehalfspacing
\maketitle\tableofcontents\newpage

@@

\section{Theory}
\subsection{What this is and what this is not}
The final product in the process of creating this programming language will 
not be perfect, or in fact, be used or adopted by anyone at all, and it is not
intended to be. Creating a good, or even usable, programming language 
implementation by today's  standards means to create an entire ecosystem of 
tools, not only including a compiler or interpreter, but also a package manager
for dealing with dependencies and a platform for people to share their work.

@@

\subsection{Different types of interpreters}
In computing, there are several types of interpreters that behave differently.
Here are some examples and their use-cases:
\begin{itemize}
    \item \emph{Just in time Compiler.} JIT-compilers blur the line between
        classical interpreters and compilers. As the name suggests, code is
        being compiled at run time of the program. The JIT-compiler can optimize
        compilation of more-often used code and therefore adapt itself to
        behave somewhat close to optimally for a given use-case. Today, this 
        technique is often used instead of classical interpreters, as
        JIT-compilers come with only advantages. The most popular example of
        a JIT-compiler would be Google's V8\footcite{V8} JavaScript engine used
        in NodeJS and any Chromium browsers.
    \item \emph{Bytecode Interpreter.} Such interpret the bytecode that was
        output by a bytecode compiler from a given piece of code. The bytecode
        is an intermediate representation of the program, used to speed up
        both compilation and interpretation. Such an intermediary bytecode is
        useful for ensuring platform independence or portability, as it will be
        the same across any machine, the only difference being the interpreter.
        The most popular example would be the Java Virtual Machine\footcite{jvm},
        a bytecode interpreter for java bytecode, also used by many other languages
        such as Kotlin, Scala or Groovy.
    \item \emph{Abstract Syntax Tree interpreter.} Source code can be
        transformed into an AST or parse tree by a parser. A simple example
        can be found under 
        \hyperref[simple interpreter]{'Implementation/The Art of Abstraction'}
        on page \pageref{simple interpreter}. Compared to bytecode interpreters
        there is a large time overhead when performing syntax related computation
        or visiting tree nodes recursively.
    \item \emph{Self interpreter.} These are interpreters that interpret the
        language they themselves are written in. This requires a program
        written in another language running an interpreter for the wanted
        language, in which runs another interpreted for and written in said
        language. Using a host language to initiate such system is called
        bootstrapping.

        Self-interpreters are closely related to self-hosting compilers. These
        are in turn compilers which are written in the language they compile.
        An example of a compiler written in itself is rustc, the Rust
        compiler\footcite{rustc}, which was bootstrapped by the OCaml language.
\end{itemize}
This paper will implement a mixture form of an AST compiler and bytecode
interpreter in-one.

\section{Implementation}
\subsection{Used Software}
At the heart of creating a program in C stands the C compiler. This project uses
\emph{gcc}, the GNU Compiler Collection, which, as the name suggests, also 
supports different languages than C. Substantiating the development process is 
\emph{GNU Make}, a general-purpose build system. It handles tasks like building
the executable step-by-step from the project source code and directory clean-up.
The C ecosystem is very easy to use and fundamentally supported by 
\emph{GNU+Linux}, the operating system used to program on.
The main code editor used was \emph{Microsoft Visual Studio Code}, providing great
features like built-in git integration.
\emph{Git} is the project management and tracking software used to back up and share
the project files on \emph{GitHub.com}, a Microsoft hosted git repository server to
store and collaborate on source code.

\subsection{The Art of Abstraction}
When dealing with problems in general, a helpful method of going about can be
to create an abstraction. This is also the case when building a programming
language interpreter. Like that, an interpreter is often times subdivided into several
components, each solving a discrete problem. One way to define
such components could be the following:
\begin{enumerate}\pagelabel{simple interpreter}
    \item Create meaningful groups of characters from the input string. This
        process is called tokenizing and is done with the component called
        tokenizer or lexer. As an example, the string 'aa+  bb' could result
        in three tokens, namely 'aa', '+' and 'bb', in case space characters
        are considered unimportant.
    \item Ordering the tokens into a so-called parse-tree or syntax-tree.
        Creating an order is important, as some calculations depend on others, e.g. multiplication
        is always done before addition. % TODO:example
    \item Technically, the parse tree could now be interpreted directly, and
        an output could be produced, as the interpreter know knows what to do
        and in what order to perform computations.
\end{enumerate}
As seen above, these components, or stages as they will be called in this paper,
each input and output data, where the output of one stage is the input of the
respective next stage.

@@

\newpage\subsubsection{File Structure}
Here is a file tree overview of the project which will be referenced
several times in the following sections of this paper. In the depicted file tree
only source files are shown, no build files or executables.
\begin{flushleft}
    \small\setlength{\DTbaselineskip}{10pt}
    \dirtree{%
        .1 project root.
        .2 .gitignore.
        .2 README.md.
        .2 LICENSE.
        .2 Makefile.
        .2 mapop.gperf.
        .2 paper.tex.
        .2 bibliography.bib.
        .2 pseudocode.
            .3 src/interpreter/parser.c.
        .2 src.
            .3 main.c.
            .3 interpreter.
                .4 bytecode.c.
                .4 bytecode.h.
                .4 error.
                    .5 error.c.
                    .5 error.h.
                .4 interpreter.c.
                .4 interpreter.h.
                .4 libraries.
                    .5 libraries.c.
                    .5 libraries.h.
                    .5 stdlib.
                        .6 stdlib.c.
                        .6 stdlib.h.
                .4 localizer.c.
                .4 localizer.h.
                .4 mappings.
                    .5 operations.c.
                    .5 operations.h.
                    .5 mapop.c.
                    .5 mapop.h.
                .4 memory.
                    .5 hashtable.c.
                    .5 hashtable.h.
                    .5 array.c.
                    .5 array.h.
                    .5 stack.c.
                    .5 stack.h.
                .4 parser.c.
                .4 parser.h.
                .4 processing.
                    .5 implementations.c.
                    .5 implementations.h.
                .4 runtime.c.
                .4 runtime.h.
                .4 tokenizer.c.
                .4 tokenizer.h.
    }
\end{flushleft}

As visible in the file tree, most of the source-code files, ending in .c,
have a corresponding header file ending in .h. This is due to the nature
of the compilation process of the C language, which is the host language
the interpreter of \name is written in.
Each header file contains only information about functions, types or variables
available to other source or header files. The actual implementations or values
are stored in the respective .c files.
% TODO:explain C compilation or link example

The pseudocode directory contains only examples used in this paper, no acutal,
working code. The paper is written in \LaTeX and BibLaTeX and compiled with
the tools pdflatex and biber. The paper's source files are paper.tex and
bibliography.bib.

The file mapop.gperf is a file to generate a perfect-hash-function\footcite{hash}
with the GNU gperf tool. The generated mapop.c file is used to map string operators
like '==' or '+' to a simpler number representation in constant time.

@@

For those who are further interested in the C programming language, there are
many internet articles or tutorials about reading and understanding or
writing and compiling C code. One such example is the article 'The C Beginner's
Handbook' provided by freeCodeCamp.org linked in the sources\footcite{freeCodeCamp}
on page \pageref{bibliography}.

@@

\subsubsection{The Parser}
The parser, implemented in src/interpreter/parser.c, generates a syntax-tree
from tokens. The algorithm used is inspired by the Pratt Parsing\footcite{pratt}
algorithm first described by Vaughan Pratt in 1973, which is a kind of precedence
parser based on recursive descent. There is a great simple explanation of the algorithm
by Jonathan Apodaca on dev.io\footcite{devio} linked in the references on page 
\pageref{bibliography} which this paper's parser takes inspiration from.

@@

To generate such a sequence of bytes, \name recursively reads through the
binary tree created by the parser and pushes found instructions onto a stack,
which is in this case used similarly to a variable-sized list.
The algorithm used is inspired by Edsger Dijkstra's 'Shunting yard 
algorithm'\footcite{shunting_yard}, traversing the tree post-order to create
a prefix notation.

@@

% TODO: add examples in /examples/
\name is built upon a well structured implementation of a recursive Pratt
parser and a basic tokenizer. Most actual features are reliant upon this
foundation, such as operators and the standard library built into \name.

@@

Following is a piece of example \name code, printing out the first ten numbers
of the fibonacci sequence, as well as a disection of how that same code is interpreted.
\code{examples/fibonacci.nts}{1}{999}
\begin{itemize}
\item The source code begins by declaring two variables \emph{a} and \emph{b}, which
    will be utilized to keep track of the data used to calculate fibonacci numbers.
    The memory allocation is done by creating a tuple of \emph{a} and \emph{b}, as suggested by 
    the comma seperating the two. Since both variables \emph{a} and \emph{b} are not yet
    known to the interpreter, they are automatically created and assigned a value
    which has no content and the \emph{undefined} typing.
\item The two variables \emph{a} and \emph{b} are then reassigned new values on line
    2 of the program. In this case it is numbers with values zero and one respectively.
    In this example, the two expressions are seperated by a space, but they can
    also be separated by newline-characters or semicolons.
\item Another variable \emph{count} is declared and defined as number in one step on line 4.
    This variable will be used to limit the amount of numbers the program outputs.
\item Yet another variable, \emph{f}, is declared and assigned a function as its
    value on line 6. In \name, normal functions do not have any arguments, 
    but they can interact with variables that were created outside their scope.
\item Lines 7 to 10 hold mathematical expressions. At first, a is assigned its current value
    plus the current value of b. Then in line 8, b is assigned the now new value of a minus the
    current value of b, which corresponds to the value of a before it was just reassigned.
    Lastly the variable \emph{count} is decremented by one.
\item Now come the interesting parts of this program.
    On lines 12 and 13, the function \emph{std.print} is referenced. This function is implemented
    in plain C code, which internally is not handled the same as regular functions in \name,
    and is wrapper to C's function \emph{printf}. The \emph{std.print} function
    takes any \name variable or value and prints it to standard output. On line 13, the "\n"
    is replaced by a newline-character, as \name can interpret Java-like escape-sequences.
\item Another standard-library function is \emph{std.if}. It takes two or three arguments.
    If the first argument is true, then the second, else the the third argument or 
    \emph{undefined} is returned. A value is true, only if it is a number not equal to zero
    or a non-empty string. In the case of this example, \emph{std.if} returns \emph{f} if 
    count is greater than zero, and an empty function if count reaches zero. The two
    brackets following the call of the \emph{std.if} function signal yet another call to the
    returned value.
\item The final thing left to do is to get the ball roling by calling the function \emph{f} for
    the first time. As stated above, it will now print numbers and recursively call itsself,
    until the variable \emph{count} reaches a value of zero.
\end{itemize}

@@

Pretty much any good, modern programming language does not only consist
of it's specification and a compiler or interpreter, but of many more
tools surrounding it.
Such tools often include a package-manager, with which people around the world
can share code. Examples would be npm\footcite{npm} or cargo\footcite{cargo}, 
the 'Node Package Manager' for JavaScript applications and and Rust's package 
manager respectively.

@@

\subsubsection{Foreign Function Interfaces}
A foreign function interface\footcite{FFI} often abbreviated as FFI,
is a mechanism by which one programming language can call functions written 
in another. This makes a language more attractive, as it can be used with
existing libraries from another language, instead of functions having to be
re-written. Enabling C function calls often allows a certain language to
directly interact with system libraries or the operating system itsself.

Foreign function interfaces are available in many modern languages.
Examples are Java, Python and Rust.
Java enables users to call C, C++ and even assembly code with the Java Native
Interface\footcite{JNI}, JNI for short, whereas within Python C function can
be run with the standard ctypes\footcite{ctypes} library.

Foreign function interfaces are specially difficult to implement, or implement
right. Let us assume a programming language interpreter \name written in C that
would like to call some functions that are part of yet another C program or
library. The problem lies in the fact that at runtime, C does not know
structure of the functions to be called, say return- or argument-types, which
is by the way where the term foreign functions originates from. This is why the
\name developer would have to provide that information by re-declaring the
function inside \name, so that, with some trickery, data can be passed to and
received from the function in the correct format, preventing a segmentation
fault. The libffi\footcite{libffi} C-library, which is also used by Java and
Python for this purpose, can be used to load and call such foreign functions.

\subsubsection{Reflection}
Reflection\footcite{reflection}, also called reflective programming, describes
the ability of a program to examine and modify its own structure and behavior 
at runtime. An example would be self-modifying code. This can easily be
achieved in assembly language, which inherently does not differentiate between
data and instructions.
Another example is Java, providing methods and classes in the java.lang.reflect
package that enable developers to examine and change properties or functions
of objects.
The advantage of reflection over not having it is that it allows for non-static
programs and program flows, effectively making smaller executables and more
resource efficient programs, as those programs can adjust themselves to run
differently under different conditions at runtime.
The easiest and simplest way to go about implementing Reflection in \name is
certainly to allow the interpretation of strings at runtime. That way, strings
holding program instructions can be created and modified and later executed.

@@

\newpage\printbibliography[heading=bibintoc]\pagelabel{bibliography}

>>>>

\documentclass[12pt,a4paper]{report}
\usepackage{listings,xcolor,setspace,hyperref,dirtree}
\usepackage{catchfile,pgffor}
\usepackage{times} % Use Times New Roman font
\usepackage[style=alphabetic,backend=biber]{biblatex}

@@

\author{Florian Donnelly}
\title{The Creation of a Programming Language}

@@

% Cite text
\newcommand{\cte}[1] {
    \cite{#1}
}

@@

% Set line spacing
\onehalfspacing

@@

\maketitle\tableofcontents\newpage

\abstract{
    This is an abstract...
}

@@

\section{Theory}
\subsection{What this is and what this is not}
The programming language and the tools that surround it are not meant to be
viable options for real-world projects. Instead, the final product, \name,
should be a programming language interpreter which includes techniques and
algorithms already present in other, well known interpreters.
The goal of this paper is to further understand the inner workings of
interpreters, transforming a human-readable programming language
into a computer-executable program.

Modern programming languages often times include additional tools, aside from
the interpreter or compiler. This includes dependency management tools to
easily create projects exdending upon existing ones, or a web-plattform for
people to share their code.

@@

\subsection{Different types of interpreters}
In computing, there are several types of interpreters that behave differently.
Here are some examples and their use-cases:
\begin{itemize}
    \item \emph{Just in time Compiler.} JIT-compilers blur the line between
        classical interpreters and compilers. As the name suggests, code is
        being compiled at run time of the program. The JIT-compiler can optimize
        compilation of more-often used code and therefore adapt itself to
        behave somewhat close to optimally for a given use-case. Today, this 
        technique is often used instead of classical interpreters, as
        JIT-compilers come with only advantages. The most popular example of
        a JIT-compiler would be Google's V8\cte{V8} JavaScript engine used
        in NodeJS and any Chromium browsers.
    \item \emph{Bytecode Interpreter.} Such interpret the bytecode that was
        output by a bytecode compiler from a given piece of code. The bytecode
        is an intermediate representation of the program, used to speed up
        both compilation and interpretation. Such an intermediary bytecode is
        useful for ensuring platform independence or portability, as it will be
        the same across any machine, the only difference being the interpreter.
        The most popular example would be the Java Virtual Machine\cte{jvm},
        a bytecode interpreter for java bytecode, also used by many other languages
        such as Kotlin, Scala or Groovy.
    \item \emph{Abstract Syntax Tree interpreter.} Source code can be
        transformed into an AST or parse tree by a parser. A simple example
        can be found under 
        \hyperref[simple interpreter]{'Implementation/The Art of Abstraction'}
        on page \pageref{simple interpreter}. Compared to bytecode interpreters
        there is a large time overhead when performing syntax related computation
        or visiting tree nodes recursively.
    \item \emph{Self interpreter.} These are interpreters that interpret the
        language they themselves are written in. This requires a program
        written in another language running an interpreter for the wanted
        language, in which runs another interpreted for and written in said
        language. Using a host language to initiate such system is called
        bootstrapping.

        Self-interpreters are closely related to self-hosting compilers. These
        are in turn compilers which are written in the language they compile.
        An example of a compiler written in itself is rustc, the Rust
        compiler\cte{rustc}, which was bootstrapped by the OCaml language.
\end{itemize}
This paper will implement a mixture form of an AST compiler and bytecode
interpreter. This is not practical in terms of improving \name, but serves
the purpose of combining different features of different interpreters into one
and therefore extending the scope of the project.

\section{Implementation}
\subsection{Used Software}
At the heart of making any project in the C programming language
stands the C compiler. This project uses
\emph{gcc}, the GNU Compiler Collection, which, as the name suggests, also 
supports more languages than solely C. Substantiating the development process is 
\emph{GNU Make}, a general-purpose build system. It handles tasks like building
the final executable program step-by-step from the project source code and directory clean-up.
The C ecosystem is very easy to use on and fundamentally supported by 
\emph{GNU+Linux}, the operating system used to program on.
The code editor used was \emph{Microsoft Visual Studio Code}, providing great
help with built-in git integration and auto-completion features.
\emph{Git} is the project management and tracking software used to back up and share
the project files on \emph{GitHub.com}, a Microsoft hosted git repository and
web front-end to store and collaborate on source code.

\subsection{The Art of Abstraction}
When dealing with problems in general, it is helpful to map out and think about
the way of approaching the problem first. A common approach is to distill the
problem situation down to a simpler one and extending the solution to take into
account all important aspects later.

The problem to solve in this paper is concisely to make a computer understand
human-written and -readable text, called source code, and follow the instructions
it proposes.
Most interpreters follow a plain scheme to go about this problem, but the
actual implementation of each of the steps can be heavily customized.
Here is the problem, broken down into three steps as commonly done in other
popular programming languages:
\begin{enumerate}\pagelabel{simple interpreter}
    \item To create meaningful groups of characters from the input string. This
        process is called 'tokenization' and is done with the component called
        tokenizer or lexer. As an example, the string 'aa + bb' could result
        in three tokens, namely 'aa', '+' and 'bb'. In this case, space characters
        are ignored. 
        This is only an example and the grouping of characters could be done 
        in any thinkable, logical way.
    \item To order the tokens into parse-tree, also called syntax-tree, as
        specified by rules. These rules are called a grammar.
        Creating an order is important, as some calculations depend on others, e.g. multiplication
        is always done before addition. An IT data structure that can accommodate
        such requirements is the binary tree. A binary tree consists of tree-nodes
        that each have a maximum of two other tree nodes as their children, and
        a tree node whose child they are, the parent node.
    \item To interpret the parse tree, traversing through it and following the
        instructions each of the tree-nodes represent. The outcome
        will be the result of the computation the program describes.
\end{enumerate}

As seen above, these components, or stages as they will be called in this paper,
each input and output data, where the output of one stage is the input of the
respective next stage.

@@

\subsubsection{File Structure}
Here is a file tree overview of the project which will be referenced
several times in the following sections of this paper. In the depicted file tree
only source files are shown. No build files, executable files or files that
are of no further importance are depicted.
\begin{flushleft}
    \scriptsize\singlespacing\setlength{\DTbaselineskip}{9pt}
    \dirtree{%
        .1 project root.
        .2 .gitignore.
        .2 README.md.
        .2 Makefile.
        .2 mapop.gperf.
        .2 pseudocode.
            .3 src/interpreter/parser.c.
        .2 src.
            .3 main.c.
            .3 interpreter.
                .4 bytecode.c.
                .4 bytecode.h.
                .4 error.
                    .5 error.c.
                    .5 error.h.
                .4 interpreter.c.
                .4 interpreter.h.
                .4 libraries.
                    .5 libraries.c.
                    .5 libraries.h.
                    .5 stdlib.
                        .6 stdlib.c.
                        .6 stdlib.h.
                .4 localizer.c.
                .4 localizer.h.
                .4 mappings.
                    .5 operations.c.
                    .5 operations.h.
                    .5 mapop.c.
                    .5 mapop.h.
                .4 memory.
                    .5 hashtable.c.
                    .5 hashtable.h.
                    .5 array.c.
                    .5 array.h.
                    .5 stack.c.
                    .5 stack.h.
                .4 parser.c.
                .4 parser.h.
                .4 processing.
                    .5 implementations.c.
                    .5 implementations.h.
                .4 runtime.c.
                .4 runtime.h.
                .4 tokenizer.c.
                .4 tokenizer.h.
    }
\end{flushleft}

As visible in the file tree, most of the source-code files, ending in .c,
have a corresponding header file ending in .h. This is due to the nature
of the compilation process of the C language\cte{c_compilation}, which is the host language
the \name interpreter is written in.
Each header file contains only information about functions, types or variables
available to other source or header files. The actual implementations or 
variable values are stored in the respective .c files.
% TODO:explain C compilation or link example

The file mapop.gperf is also a source file. It is used in conjunction with the GNU gperf
tool. The tool creates a perfect hash function\cte{hash}, which is output to the file
src/interpreter/mappings/mapop.c. The perfect hash function is used to translate
single- or multi-character symbols, such as '!=' and '+', to their respective
interpreter internal representation in constant time.

@@

For those who are further interested in the C programming language, there are
many internet articles or tutorials about reading and understanding or
writing and compiling C code. One such example is the article 'The C Beginner's
Handbook' provided by freeCodeCamp.org linked in the sources\cte{freeCodeCamp}
on page \pageref{bibliography}.

@@

\subsubsection{The Parser}
The parser, implemented in src/interpreter/parser.c, generates a syntax-tree
from tokens. The algorithm used is inspired by the Pratt Parsing\cte{pratt}
algorithm first described by Vaughan Pratt in 1973, which is a kind of precedence
parser based on recursive descent. There is a great simple explanation of the algorithm
by Jonathan Apodaca on dev.io\cte{devio} linked in the references on page 
\pageref{bibliography} which this paper's parser takes inspiration from.

@@

To generate such a sequence of bytes, \name recursively reads through the
binary tree created by the parser and pushes found instructions onto a stack,
which is in this case used similarly to a variable-sized list.
The algorithm used is inspired by Edsger Dijkstra's 'Shunting yard 
algorithm'\cte{shunting_yard}, traversing the tree post-order to create
a prefix notation.

@@

% TODO: add examples in /examples/
\name is built upon the implementation of a recursive Pratt
parser and a basic tokenizer. Most actual features are reliant upon this
foundation, such as operators and the standard library built into \name.

@@

Following is a piece of example \name code, printing out the first ten numbers
of the fibonacci sequence, as well as a disection of how that same code is interpreted.
\code{examples/fibonacci.nts}{1}{999}
\begin{itemize}
\item The source code begins by declaring two variables \emph{a} and \emph{b}, which
    will be utilized to keep track of the data used to calculate fibonacci numbers.
    The memory allocation is done by creating a tuple of \emph{a} and \emph{b}, as suggested by 
    the comma seperating the two. Since both variables \emph{a} and \emph{b} are not yet
    known to the interpreter, they are automatically created and assigned a value
    which has no content and the \emph{undefined} typing.
\item The two variables \emph{a} and \emph{b} are then reassigned new values on line
    2 of the program. In this case it is numbers with values zero and one respectively.
    In this example, the two expressions are seperated by a space, but they can
    also be separated by newline-characters or semicolons.
\item Another variable \emph{count} is declared and defined as number in one step on line 4.
    This variable will be used to limit the amount of numbers the program outputs.
\item Yet another variable, \emph{f}, is declared and assigned a function as its
    value on line 6. In \name, normal functions do not have any arguments, 
    but they can interact with variables that were created outside their scope.
\item Lines 7 to 10 hold mathematical expressions. At first, a is assigned its current value
    plus the current value of b. Then in line 8, b is assigned the now new value of a minus the
    current value of b, which corresponds to the value of a before it was just reassigned.
    Lastly the variable \emph{count} is decremented by one.
\item Now come the interesting parts of this program.
    On lines 12 and 13, the function \emph{std.print} is referenced. This function is implemented
    in plain C code, which internally is not handled the same as regular functions in \name,
    and is wrapper to C's function \emph{printf}. The \emph{std.print} function
    takes any \name variable or value and prints it to standard output. On line 13, the "\\n"
    is replaced by a newline-character, as \name can interpret Java-like escape-sequences.
\item Another standard-library function is \emph{std.if}. It takes two or three arguments.
    If the first argument is true, then the second, else the the third argument or 
    \emph{undefined} is returned. A value is true, only if it is a number not equal to zero
    or a non-empty string. In the case of this example, \emph{std.if} returns \emph{f} if 
    count is greater than zero, and an empty function if count reaches zero. The two
    brackets following the call of the \emph{std.if} function signal yet another call to the
    returned value.
\item The final thing left to do is to get the ball roling by calling the function \emph{f} for
    the first time. As stated above, it will now print numbers and recursively call itsself,
    until the variable \emph{count} reaches a value of zero.
\end{itemize}

@@

Pretty much any good, modern programming language does not only consist
of it's specification and a compiler or interpreter, but of many more
tools surrounding it.
Such tools often include a package-manager, with which people around the world
can share code. Examples would be npm\cte{npm} or cargo\cte{cargo}, 
the 'Node Package Manager' for JavaScript applications and and Rust's package 
manager respectively.

@@

\subsubsection{Foreign Function Interfaces}
A foreign function interface\cte{FFI} often abbreviated as FFI,
is a mechanism by which one programming language can call functions written 
in another. This makes a language more attractive, as it can be used with
existing libraries from another language, instead of functions having to be
re-written. Enabling C function calls often allows a certain language to
directly interact with system libraries or the operating system itsself.

Foreign function interfaces are available in many modern languages.
Examples are Java, Python and Rust.
Java enables users to call C, C++ and even assembly code with the Java Native
Interface\cte{JNI}, JNI for short, whereas within Python C function can
be run with the standard ctypes\cte{ctypes} library.

Foreign function interfaces are specially difficult to implement, or implement
right. Let us assume a programming language interpreter \name written in C that
would like to call some functions that are part of yet another C program or
library. The problem lies in the fact that at runtime, C does not know
structure of the functions to be called, say return- or argument-types, which
is by the way where the term foreign functions originates from. This is why the
\name developer would have to provide that information by re-declaring the
function inside \name, so that, with some trickery, data can be passed to and
received from the function in the correct format, preventing a segmentation
fault. The libffi\cte{libffi} C-library, which is also used by Java and
Python for this purpose, can be used to load and call such foreign functions.

\subsubsection{Reflection}
Reflection\cte{reflection}, also called reflective programming, describes
the ability of a program to examine and modify its own structure and behavior 
at runtime. An example would be self-modifying code. This can easily be
achieved in assembly language, which inherently does not differentiate between
data and instructions.
Another example is Java, providing methods and classes in the java.lang.reflect
package that enable developers to examine and change properties or functions
of objects.
The advantage of reflection over not having it is that it allows for non-static
programs and program flows, effectively making smaller executables and more
resource efficient programs, as those programs can adjust themselves to run
differently under different conditions at runtime.
The easiest and simplest way to go about implementing Reflection in \name is
certainly to allow the interpretation of strings at runtime. That way, strings
holding program instructions can be created and modified and later executed.

@@

\newpage\section{Bibliography}
\pagelabel{bibliography}\printbibliography[heading=none]
